Real-Time Certified Probabilistic Pedestrian Forecasting

### 2017-02-16 pedestrian_forecasting_iros_2017.tex
editing through implementation

\begin{abstract}
That autonomous vehicles will come to dominate our streets is imminent.
This motivates the need for a real-time probabilistic forecasting algorithm for pedestrians, cyclists, and other agents,
as it forms a necessary step in assessing the risk (and therefore the cost) we should expect to incur.
In this paper, we present a novel approach to probabilistic forecasting for pedestrians based on weighted sums of ordinary differential equations.
The resulting algorithm is embarrassingly parallel, and trained on historical trajectory information within a fixed scene.
When compared with MDP-based methods, our algorithm appears to be superior from the standpoint of precision and recall.
\end{abstract}

@@

\section{Model}\label{sec:model}
This paper's goal is to generate a time-dependent probability density over $\mathbb{R}^2$, which predicts the true location of an agent in the future.
The input to the algorithm at runtime is a noisy measurement of position and velocity, $\hat{x}_0, \hat{v}_0 \in \mathbb{R}^2$.
If the (unknown) location of agent at time $t$ is given by $x_t \in \mathbb{R}^2$, then the distribution we seek is the posterior $\rho_t(x_t) := \Pr( x_t \mid \hat{x}_0, \hat{v}_0 )$.

@@

\subsection{The Variables of the Model}
The model concerns the position of an agent $x_t \in \mathbb{R}^2$ for $t \in [0,T]$ for some $T>0$.
We denote the position and velocity at time $t=0$ by $x_0$ and $v_0$ respectively.
At $t=0$, we obtain a measurement of position and velocity, denotes by $\hat{x}_0$ and $\hat{v}_0$.
Lastly, we have a variety of motion models, parametrized by a set $\mathcal{M}$ (described in the sequel).
For each model $m \in \mathcal{M}$, a trajectory $\check{x}_t$ given the initial position and velocity $x_0$ and $v_0$.
All these variables are probabilistically related to one another in a (sparse) Bayesian network, which we will describe next.

@@

This is a powerful result, because it allows us to solve for $\Pr( x_t, \hat{x}_0, \hat{v}_0)$  (and thus $\rho_t(x_t)$) at times $t=\Delta t, \dots, N_t \Delta t$ by a single computation of $\Phi_{k,1}^{\ell \Delta t \bar{s}}( x_0^{i,j})$ for each $\ell \in  \{-N_t,\dots,N_t\}$, $k \in \{1,\dots,n\}$, and $x_0^{i,j} \in \Gamma_L( \hat{x}_0 , N_x)$.
To explain this, assume we are given $\Phi_{k,1}^{m \Delta t \bar{s}}(x_0^{i,j})$ for all $|m| \leq \ell$.
We can then use the regular partition
\begin{align}
	\mathcal{P}_\ell := \left\{ \frac{m}{\ell} \bar{s} \right\}_{m=-\ell}^{\ell} \label{eq:partition}
\end{align}
of $[-\bar{s}, \bar{s}]$ to compute a Riemann sum approximation using Theorem \ref{thm:main}.
The partition $\mathcal{P}_\ell$ has a width of size, $\Delta s = \bar{s} / \ell$, and substitution in \eqref{eq:approximation 3} yields an approximation of $\Pr( x_t, \hat{x}_0, \hat{v}_0)$ given by
\begin{align*}
	\begin{split}
		&\frac{\bar{s}}{\ell} \sum_{i,j,k,m} W(k,m \bar{s} / \ell ,i,j,\hat{x}_0) \left. \Pr( x_t \mid \check{x}_t ) \right|_{ \check{x}_t = \Phi_{k,1}^{m \bar{s} \Delta t}( x_0^{i,j}) } \\
		&\qquad+\Pr( x_t , lin, \hat{x}_0, \hat{v}_0 ).
	\end{split}
\end{align*}
As we've already computed $\Phi_{k,1}^{m \bar{s} \Delta t}$ for $| m | \leq \ell$ by assumption, the only obstacle to computing this sum is the computation of the weights $W(k,m \bar{s} / \ell ,i,j,\hat{x}_0)$.
%Admittedly, this means our approximations might be imprecise over very short times, but more accurate over longer time-horizons.
The procedure to compute  $\rho_t(x_t)$ is summarized in Algorithm \ref{alg:1}.

\begin{algorithm}[t]
		  \caption{Algorithm to Compute  $\rho_t$ at times $t= \Delta t, 2 \Delta t, \dots, N_t \Delta t$.}
		  \label{alg:1}
		  \begin{algorithmic}[1]
		    \Require $\bar{s} > 0, N_t \in \mathbb{N}, \Delta t > 0,   \{ X_k \}_{k=1}^{n}, \hat{x}_0, \hat{v}_0, \Gamma_L( \hat{x}_0, N_x)$,
		    routines for $\Pr(x_0 \mid M), \Pr(x_0 \mid \hat{x}_0), \Pr(v_0 \mid \hat{v}_0 ), \Pr(M \in \mathcal{M}), \Pr(s \mid k)$ .
		    \For {$\ell = 1,\dots,N_t$}
		    \State Compute and store $\Phi_{k,1}^{ \pm \bar{s} \ell \Delta t}( x_0^{i,j} )$ for all $x_0^{i,j} \in \Gamma_L( \hat{x}_0 ; N_x)$.
		    \For {$m = -\ell,\dots,\ell$}
		    \State Compute $W(k,s,i,j,\hat{x}_0)$ for $s = \bar{s} m / \ell$ using \eqref{eq:pgm}.
		    \EndFor  
		    \State Construct a $\mathcal{O}( \Delta x + \bar{s}/\ell + \varepsilon_{\rm tol})$ approximation of $\Pr(\check{x}_t, \hat{x}_0, \hat{v}_0)$ at $t=\ell \Delta t$ via Theorem \ref{thm:main} 
		    with partition $\mathcal{P}_\ell$ of \eqref{eq:partition}.
		   \label{alg1:step2}.
		    \State Normalize with respect to $\check{x}_t$ to obtain an approximation of $\Pr(\check{x}_t \mid \hat{x}_0, \hat{v}_0)$.
		    \State Convolve with respect to $\Pr( x_t \mid \check{x}_t)$ to obtain $\rho_t(x_t)$ at $t= \ell \cdot \Delta t$
		    \EndFor
		  \end{algorithmic}
\end{algorithm}

@@

%Steps \ref{alg1:step1} and \ref{alg1:step2} are where the bulk of the work occurs.
%Fortunately, these two steps are embarrassingly parallel and yield tolerable computational complexity for real-time computation.
For fixed $k$ $i$ and $j$, the computation of $\Phi_{k,1}^t(x_0^{i,j})$ at each $t = \{-N_t\bar{s} ,\dots, N_t \bar{s}\}$ takes $\mathcal{O}(N_t)$ time using an explicit finite difference scheme and can be done in parallel for each $k \in \{1,\dots, n\}$ and $x_0^{i,j} \in \Gamma_L( \hat{x}_0; N_x)$.
Similarly, computing $W(k,s,i,j,\hat{x}_0)$ constitutes a series of parallel function evaluations over tuples $(k,s,i,j)$ of the posterior distributions described in \eqref{eq:pgm}, where the continuous variable $s$ is only required at finitely many places in Algorithm \ref{alg:1}.
If the posteriors represented by the arrows in \eqref{eq:pgm} are efficiently computably then the computation of $W(k,s,i,j, \hat{x}_0)$ is equally efficient.
%We summarize this in the following theorem:
%\begin{thm}
%	Suppose we have an accurate scheme for integrating ODEs in a computation time proportional to the integration time-interval\footnote{For example, 4th order explicit Runge-Kutta}.  
%	Algorithm \ref{alg:main} yields an $\mathcal{O}( \Delta x + \bar{s} / \ell + \epsilon_{\rm tol})$ accurate approximation of $\rho_t$ in $\mathcal{O}( N_t n N_x^2 / N_p )$ computation time given $N_p$ processing units and $n$ vector fields.
%\end{thm}

@@

This section establishes our methods performance and compares it to the model from \cite{Kitani2012} and a random walk.
We implement our model as well as our evaluation code in Python 2.6. 
Our implementation is available online\footnote{https://github.com/OkayHughes/iros2017\_pedestrian\_forecasting}.
Our test data used the bicycle annotations in \cite{Robicquet2016}. 
We did a 2-fold cross validation on 4 different scenes of varying complexity. 
Our analysis used the same partitions of the trajectories for all three predictors.

@@

\begin{table}
	\begin{center}
		\caption{Comparison of runtimes of the various algorithms.}
		\label{tab:time}
		\renewcommand{\arraystretch}{1.5}% Spread rows out...
		\begin{tabular}{||c | c c  c ||} 
			\hline
			& Our Predictor & Random Walk & Kitani et al. \\ [0.5ex] 
			\hline 
			$\frac{\mathrm{time}}{\mathrm{frame}}$ & 0.0169s & 2.1E-7s & 0.0614s \\
			\hline
		\end{tabular}
	\end{center}
\end{table}

@@

\section{Conclusion} 
\label{sec:conclusion}
This paper presents a real-time approach to probabilistic pedestrian modeling. We demonstrate the ability to accurately predict the final location of a pedestrian with superior accuracy compared to a state-of-the-art with large additional contribtions in three critical areas: 1) prediction speed is highly efficient enabling tight loop real-time control;  2) our prediction uncertainty bound much more accurately reflect potential paths for the human limiting false negatives and maximize the utility and safety of the predictions; and 3) the prediction made by the proposed approach do not require knowledge of the end point  making them applicable to a much wider variety of real world applications. These three features have great promise for robotic applications and we see the integration of such techniques with autonomous system control as one pathway to enable safer operations in shared human-robot spaces. For future directions we plan to optimize the naive python implementation using GPU parallelization to further increase speed and ease integration into system's control loop.

>>>>

\begin{abstract}
The success of autonomous systems will depend upon their ability to safely navigate human-centric environments.
This motivates the need for a real-time, probabilistic forecasting algorithm for pedestrians, cyclists, and other agents since these predictions will form a necessary step in assessing the risk of any action.
This paper presents a novel approach to probabilistic forecasting for pedestrians based on weighted sums of ordinary differential equations that are learned from historical trajectory information within a fixed scene.
The resulting algorithm is embarrassingly parallel and is able to work at real-time speeds using a naive Python implementation.
The quality of predicted locations of agents generated by the proposed algorithm is validated on a variety of examples and considerably higher than existing state of the art approaches over long time horizons.
%When compared with MDP-based methods, this algorithm 
\end{abstract}

@@

\section{Model}\label{sec:model}
This paper's goal is to generate a time-dependent probability density over $\mathbb{R}^2$, which predicts the true location of an agent in the future.
The input to the algorithm at runtime is a noisy measurement of position and velocity, $\hat{x}_0, \hat{v}_0 \in \mathbb{R}^2$.
If the (unknown) location of agent at time $t$ is given by $x_t \in \mathbb{R}^2$, then the distribution we seek is the posterior $\rho_t(x_t) := \Pr( x_t \mid \hat{x}_0, \hat{v}_0 )$ for each time $t \in \{\Delta t, \dots, N_t \Delta t \}$ for some user-specified $N_t \in \mathbb{N}$ and $\Delta t \in \mathbb{R}$.

@@

\subsection{The Variables of the Model}
The model concerns the position of an agent $x_t \in \mathbb{R}^2$ for $t \in [0,N_t \Delta t]$.
We denote the position and velocity at time $t=0$ by $x_0$ and $v_0$ respectively.
At $t=0$, we obtain a measurement of position and velocity, denotes by $\hat{x}_0$ and $\hat{v}_0$.
Lastly, we have a variety of motion models, parametrized by a set $\mathcal{M}$ (described in the sequel).
For each model $m \in \mathcal{M}$, a trajectory $\check{x}_t$ given the initial position and velocity $x_0$ and $v_0$.
All these variables are probabilistically related to one another in a (sparse) Bayesian network, which we will describe next.

@@

This is a powerful result since it allows us to compute $\Pr( x_t, \hat{x}_0, \hat{v}_0)$  (and thus $\rho_t(x_t)$) at times $t \in \{\Delta t, \dots, N_t \Delta t\}$ with a single computation of $\Phi_{k,1}^{\ell \Delta t \bar{s}}( x_0^{i,j})$ for each $\ell \in  \{-N_t,\dots,N_t\}$, $k \in \{1,\dots,n\}$, and $x_0^{i,j} \in \Gamma_L( \hat{x}_0 , N_x)$.
To appreciate this, assume we are given $\Phi_{k,1}^{m  \bar{s} \Delta t}(x_0^{i,j})$ for all $m \in \{ -\ell, \ldots, \ell \}$.
We can then use the regular partition
\begin{align}
	\mathcal{P}_\ell := \left\{ \frac{m}{\ell} \bar{s} \right\}_{m=-\ell}^{\ell} \label{eq:partition}
\end{align}
of $[-\bar{s}, \bar{s}]$ to compute a Riemann sum approximation using Theorem \ref{thm:main}.
The partition $\mathcal{P}_\ell$ has a width of size, $\Delta s = \bar{s} / \ell$, and substitution in \eqref{eq:approximation 3} yields an approximation of $\Pr( x_t, \hat{x}_0, \hat{v}_0)$ given by
\begin{align*}
	\begin{split}
		&\frac{\bar{s}}{\ell} \sum_{i,j,k,m} W(k,m \bar{s} / \ell ,i,j,\hat{x}_0) \left. \Pr( x_t \mid \check{x}_t ) \right|_{ \check{x}_t = \Phi_{k,1}^{m \bar{s} \Delta t}( x_0^{i,j}) } \\
		&\qquad+\Pr( x_t , lin, \hat{x}_0, \hat{v}_0 ).
	\end{split}
\end{align*}
As we have already computed $\Phi_{k,1}^{m \bar{s} \Delta t}$ for $| m | \leq \ell$ by assumption, the only obstacle to computing this sum is the computation of the weights $W(k,m \bar{s} / \ell ,i,j,\hat{x}_0)$.
%Admittedly, this means our approximations might be imprecise over very short times, but more accurate over longer time-horizons.
The procedure to compute  $\rho_t(x_t)$ is summarized in Algorithm \ref{alg:1}.

\begin{algorithm}[t]
		  \caption{Algorithm to compute  $\rho_t$ for each $t \in  \{\Delta t, 2 \Delta t, \dots, N_t \Delta t$\}.}
		  \label{alg:1}
		  \begin{algorithmic}[1]
		    \Require $\bar{s} > 0$, $N_t \in \mathbb{N}$, $\Delta t > 0$,  $ \{ X_k \}_{k=1}^{n}$, $\hat{x}_0$, $\hat{v}_0$, $\Gamma_L( \hat{x}_0, N_x)$, $\Pr(x_0 \mid M), \Pr(x_0 \mid \hat{x}_0), \Pr(v_0 \mid \hat{v}_0 ), \Pr(M \in \mathcal{M}), \text{ and } \Pr(s \mid k)$.
		    \For {$\ell \in \{1,\dots,N_t\}$}
		    \State Compute $\Phi_{k,1}^{ \pm \bar{s} \ell \Delta t}( x_0^{i,j} )$ for all $x_0^{i,j} \in \Gamma_L( \hat{x}_0 ; N_x)$.
		    \For {$m \in \{-\ell,\dots,\ell\}$}
		    \State Compute $W(k,s,i,j,\hat{x}_0)$ for $s = \bar{s} m / \ell$.
		    \EndFor  
		    \State Compute $\Pr(\check{x}_t, \hat{x}_0, \hat{v}_0)$ at $t=\ell \Delta t$ via Theorem \ref{thm:main} 
%		    \State Construct a $\mathcal{O}( \Delta x + \bar{s}/\ell + \varepsilon_{\rm tol})$ approximation of $\Pr(\check{x}_t, \hat{x}_0, \hat{v}_0)$ at $t=\ell \Delta t$ via Theorem \ref{thm:main} 
		    \hspace*{12pt} with partition $\mathcal{P}_\ell$ of \eqref{eq:partition}.\label{alg1:step2}
		    \State Apply Bayes' Theorem to obtain $\Pr(\check{x}_t \mid \hat{x}_0, \hat{v}_0)$.
		    \State Compute $\rho_t(x_t)$ at $t= \ell \cdot \Delta t$ via \eqref{eq:convolve}.
		    \EndFor
		  \end{algorithmic}
\end{algorithm}

@@

%Steps \ref{alg1:step1} and \ref{alg1:step2} are where the bulk of the work occurs.
%Fortunately, these two steps are embarrassingly parallel and yield tolerable computational complexity for real-time computation.
For fixed $k$, $i$, and $j$, the computation of $\Phi_{k,1}^t(x_0^{i,j})$ at each $t = \{-N_t\bar{s} ,\dots, N_t \bar{s}\}$ takes $\mathcal{O}(N_t)$ time using an explicit finite difference scheme and can be done in parallel for each $k \in \{1,\dots, n\}$ and $x_0^{i,j} \in \Gamma_L( \hat{x}_0; N_x)$.
Similarly, computing $W(k,s,i,j,\hat{x}_0)$ constitutes a series of parallel function evaluations over tuples $(k,s,i,j)$ of the posterior distributions described in \eqref{eq:pgm}, where the continuous variable $s$ is only required at finitely many places in Algorithm \ref{alg:1}.
If the posteriors represented by the arrows in \eqref{eq:pgm} are efficiently computably then the computation of $W(k,s,i,j, \hat{x}_0)$ is equally efficient.
%We summarize this in the following theorem:
%\begin{thm}
%	Suppose we have an accurate scheme for integrating ODEs in a computation time proportional to the integration time-interval\footnote{For example, 4th order explicit Runge-Kutta}.  
%	Algorithm \ref{alg:main} yields an $\mathcal{O}( \Delta x + \bar{s} / \ell + \epsilon_{\rm tol})$ accurate approximation of $\rho_t$ in $\mathcal{O}( N_t n N_x^2 / N_p )$ computation time given $N_p$ processing units and $n$ vector fields.
%\end{thm}

@@

This section establishes our methods performance and compares it to the model from \cite{Kitani2012} and a random walk.
We implement our model as well as our evaluation code in Python 2.6. 
Our implementation is available online\footnote{https://github.com/OkayHughes/iros2017\_pedestrian\_forecasting}.
Our test data used a total  annotations in \cite{Robicquet2016}. 
We did a 2-fold cross validation by using $20\%$ of the data for testing and the remainder for training.
We used the same partitions of the trajectories for all three predictors.
Our analysis was conducted on 4 different scenes of varying complexity with a total of $350$ total annotations.

@@

\begin{table}
	\begin{center}
		\caption{Comparison of runtimes of the various algorithms.}
		\label{tab:time}
		\renewcommand{\arraystretch}{1.5}% Spread rows out...
		\begin{tabular}{||c | c c  c ||} 
			\hline
			& Our Algorithm & Random Walk & Kitani et al. \\ [0.5ex] 
			\hline 
			$\frac{\mathrm{time}}{\mathrm{frame}}$ & 0.0169s & 2.1E-7s & 0.0614s \\
			\hline
		\end{tabular}
	\end{center}
\end{table}

@@

\section{Conclusion} 
\label{sec:conclusion}
This paper presents a real-time approach to probabilistic pedestrian modeling. 
We demonstrate the ability to accurately predict the final location of a pedestrian with superior accuracy compared to a state-of-the-art with additional contributions in three critical areas: 1) prediction speed is efficient enabling tight loop real-time control;  2) our error bound on the predictions accurately reflects potential paths for the human limiting false negatives and maximizing the utility and safety of the predictions; and 3) the prediction made by the proposed approach do not require knowledge of the end point  making them applicable to a wide variety of real world applications. 
These three features have great utility for robotic applications, and we see the integration of such techniques with autonomous system control as one pathway to enable safer operations in shared human-robot spaces. 
We plan to explore the optimization of the naive python implementation using GPU parallelization to further increase speed and ease integration into system's control loop.
